---
title: "Exclusions"
author: "Alex Holcombe"
date: "2/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)

```

Load data and estimates

```{r load data, echo=FALSE, message=FALSE}

#Import raw data
dataPath<- file.path("../Data/")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") ) 

resultsPath<- file.path("Results")
estimates<- read.csv(file = file.path(resultsPath,"/backwards2E2_paramEstimates.csv"))

```

Fix the variable names if not done

```{r, echo=FALSE, message=FALSE}
#Below should now be done by import
if ('target' %in% names(data)) {
  names(data)[names(data) == 'target'] <-'stream'
  names(data)[names(data) == 'condition'] <-'orientation'
  data$stream[ data$stream==1 ] <- "Left"
  data$stream[ data$stream==2 ] <- "Right"
}
if (is.numeric(data$targetSide)) {
  data$targetSide[ data$targetSide==0 ] <- "Both"
  data$targetSide[ data$targetSide==1 ] <- "Left"
  data$targetSide[ data$targetSide==2 ] <- "Right"
}
if (is.numeric(data$orientation)) {
  #mutate condition to Orientation
  data$orientation[ data$orientation==1 ] <- "Canonical"
  data$orientation[ data$orientation==2 ] <- "Inverted"
}
```

Add the parameter estimates to the tibble

```{r, echo=FALSE, message=FALSE, fig.height=360, fig.width=10}
#want fig.height of 10 per subject

condtnVariableNames <- c("subject","stream", "orientation","targetSide")  

#Add R parameter estimates to data tibble
#To merge them, unfortunately have to convert to dataframe first because dplyr has no binding of tibbles.
# https://stackoverflow.com/questions/45643886/how-to-combine-tibbles-with-common-but-non-identical-columns#45644319
#Not sure what is lost as a result.
dd<- as_data_frame(data)
es<- as_data_frame(estimates)
#Looks like each subject has only 8 efficacy values somehow, it's because there was an extra
#variable, X
es$X <- NULL
#oldNoNa<- merge(dd,es) #loses the is.na(targetSP) rows
dat<- merge(dd,es, all = TRUE)
dg<-dat %>% filter( !is.na(targetSP) ) #excluding 1-target trials rows reporting stream on opposite side. Eventually, use those to evaluate swaps.

```

Merge operation lost   targetSP==NA in data, but now fixed
```{r, echo=FALSE}
table(dat$targetSP,useNA="ifany")

#dat %>% filter( is.na(targetSP) )
```

One-target:

BE has e=.47 on left, .62 on right
BJ is just near chance and fails likelihood ratio test for 1-target, like did for 2-targets.
BO did badly for canonical on left, but passed lr test.
CF did badly for inverted on left, but passed lr test.
CL did pretty badly but got lucky, passed.
AP flunked one of the lr tests which is good because flunked some in two-target too.
So, lr test seems to be working well enough.

BG, BJ, and AP have sigma near 4, clear outliers.

## Exclusions based on two-target conditions

Notes from histsWithFitsTwoTarget.Rmd:
AP, especially BJ does flunk likelihood test! CY passes but data not horrifyingly bad.
Exclude by likelihood-ratio test.
Participant AI, BB suspicion of  attending to one side from twoTargets condition.


Exclude by likelihood ratio test and efficacy.

For each subject, calculate how many two-target streams they failed. If 3 or 4 of 4, exclude (didn't think to mention 1-target condition in prereg).


Also check whether average efficacy is below 10% in both streams in either the upright or the reversed orientation.

```{r exclusions}

twoTarget<- dg %>% filter(targetSide=="both")

#first, collapse over trials
#https://stackoverflow.com/questions/51978519/summarise-numeric-columns-return-last-value-of-non-numeric
tw<- twoTarget %>% group_by(subject,stream,orientation) %>% select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })

#for each subject, count pLRtest instances >.05 and efficay <.1
B<-tw %>% ungroup %>% group_by(subject) %>% summarise(LRtestFails = sum(pLRtest>.05), effFails = sum(efficacy<.1))                                                              
#B %>% print(n=Inf)
LRtestLosers <- B %>% filter(LRtestFails>=3)
LRtestLosers<- unique(LRtestLosers$subject)
# efficacy is below 10% in both streams in either the upright or the reversed orientation.
# To calculate that, need to group by subject and orientation
C<-tw %>% ungroup %>% group_by(subject,orientation) %>% summarise(bothEffsFailed = sum(efficacy<.1)==2)
#Find which subjects, if any, has a bothEffsFailed==TRUE
lowEff<- C %>% filter(bothEffsFailed==TRUE)
lowEffLosers <- unique(lowEff$subject)
```

## Exclusions based on one-target conditions

There are 4 one-target conditions also, upright/reversed x left/right.
Will exclude subject if failed likelihood test for both left and right in either condition.

Will also exclude if difference in efficacy left-right exceeds?
```{r }
#Calculate exclusions based on one-target performance
oneTarget<- dg %>% filter(targetSide!="both")
one<- oneTarget %>% group_by(subject,targetSide,orientation) %>% 
  select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
  select(-trial,-targetSP,-respSP,-respOrder) %>% #clear detritus so easier to see
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })

#Look at the difference in efficacy between the left and the right
effDiff <- function(df) {
  #dgi<<-df
  #STOP
  effLeft<- (df %>% filter(targetSide=="left"))$efficacy 
  effRight<- (df %>% filter(targetSide=="right"))$efficacy
  df$effDiff<- effLeft - effRight
  df$effLeft <- effLeft; df$effRight <- effRight
  
  df$efficacy<- mean(c(effLeft,effRight))
  df$effDiff<- effLeft - effRight
  
  failedBothLR <- ((df %>% filter(targetSide=="left"))$pLRtest >= .05) &
    ((df %>% filter(targetSide=="right"))$pLRtest >= .05)
  df$failedBothLR <- failedBothLR
  
  return(df[1,]) #return first row, so it does get collapsed
}

oneDiff<- one %>% group_by(subject,orientation) %>% do( effDiff(.) )
oneDiff %>% select(subject,targetSide,orientation,efficacy,effDiff,pLRtest,failedBothLR)

oneTargetLosers <- oneDiff %>% filter(failedBothLR == TRUE, effDiff > .4)
oneTargetLosers<- unique(oneTargetLosers$subject)
```

Make scatterplot of efficacy for left vs. right single target

```{r}
#To do that, do I'd either need to unmelt (spread) but rather than do that, I programmed putting effLeft and effRight on same row into effDiff
library(ggplot2)
g<- ggplot(oneDiff, aes(x=effLeft, y=effRight, color=orientation)) + geom_point()
g<-g +  geom_abline(intercept = 0, slope = 1) + theme_bw() + xlim(0,1) + ylim(0,1)
g +    geom_rug(col=rgb(.5,0,0,alpha=.2))

```

Make scatterplot of efficacy difference between left and right single target on horizontal axis and dual-target difference on vertical axis.

```{r}

```

I now have LRtestLosers, lowEffLosers, and oneTargetLosers

Actually exclude the losers? Probably better if flag them so can run inferential stats both with and without them.

```{r excludeAll, echo=FALSE}
allLosers <- unique( c(LRtestLosers, lowEffLosers, oneTargetLosers) )
#afterExclusions<- dat %>% filter(  !(subject %in% lowEffs)   )
dat <- dat %>% mutate(exclude = (subject %in% allLosers))
#table(dat$subject,dat$exclude)
```



Next stop, inferential stats?  Save data with estimates and exclude field added.

```{r exclude, echo=FALSE}
pathToSave <- file.path( "..","Data")
saveRDS(dat, file.path(pathToSave,"backwards2E2_dataWithEstimatesAndExcludeColumn.rda"), compress = "bzip2")
```
