---
title: "trialStreaks"
author: "Alex Holcombe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

For each trial (row of the tibble), find the n-1, n-2, n-3 trial etc., put its orientation into a new column called neg1orient, neg2orient, etc.  Then can filter by whether they equal the current orientation, and analyze left bias with percent correct the dependent variable.

## Implementation

One could make a new grouping column that is trialnum / 4 that would group together every 4 trials. That would analyze only every 4 trials, which would avoid the problem of correlated trials, but would not provide a running streak assessment. 

rolling window calculations, autocorrelation
from: http://www.business-science.io/timeseries-analysis/2017/08/30/tidy-timeseries-analysis-pt-4.html
"apply tq_mutate() using the lag.xts function. We can provide column names for the new columns by prefixing “lag_” to the lag numbers, k, which the sequence from 1 to 28. The output is all of the lags for each package."

## Statistical issues
Did van der Burg or any of them address statistical independence? Alais' latest paper does not, so probably they never did.

```{r load data, echo=FALSE, message=FALSE}
library(tidyverse)

#Import raw data
dataPath<- file.path("Data/")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E1_rawDataFromMAT.rda") ) 

#to work with dplyr, can't have array field like letterSeq
data$letterSeq<- NULL

#tidy data
df<- as_tibble(data)

```

I want only raw data, not estimates, but need to get exclude column from estimates to screen out excluded Ss
```{r, echo=FALSE}
library(readr)
resultsPath <- file.path("Results")
estimates<- readr::read_tsv(file=file.path(resultsPath,"backwards2E1_paramEstimatesWithExcludeColumn.tsv"))
```
Determine Ss to exclude and join to raw data
```{r, echo=FALSE}
#I don't remember what the excluded column is that's already there, but it excludes more Ss than we want
excludeOrNot <- estimates %>% group_by(subject) %>% summarise(exclude = last(exclude))

#Join with raw data
df<- as_tibble( merge(excludeOrNot,df) )
 #join
```

[This post](http://www.business-science.io/timeseries-analysis/2017/08/30/tidy-timeseries-analysis-pt-4.html) covers autocorrelation.

Demonstrate I can calculate lags with tiny subset of data
```{r}
library(tidyquant)
#tidyquant functions search for and key off a date or POSIXct column. 
#So, need to convert trial to that.
dfd <- df %>% mutate(trialDate = as_date(trial))
dfd <- dfd %>% filter(exclude==FALSE)

#remove some crud so easier to view
dfd<- dfd %>% select(-allRTs,-respSP,-resp,-targetSP,-excluded)
#Critical column (orientation) must be numeric
dfe<- dfd %>% mutate(orientCI = ifelse(orientation=="Canonical",1,-1))

k <- 1:4
col_names <- paste0("lag_", k)

lags <- dfe %>% slice(1:10) %>%
    tidyquant::tq_mutate(
        select     = orientCI,
        mutate_fun = lag.xts,
        k          = 1:4,
        col_rename = col_names
    )
print(lags %>% select(-exclude,-stream))
```

But need to do this for every condition. And want to do it after have brought together left stream and right stream correct, so can plot the autocorrelation for the left bias.
Look for where I did percent correct

Try to bring left and right stream onto same row, so can calculate left bias on individual trials. 
```{r}
library(data.table)

lr <- data.table::dcast(df, subject+trial+orientation~stream,value.var="SPE")
lr<-as_tibble(lr)
```


```{r}
lr<- lr %>% mutate(orientCI = ifelse(orientation=="Canonical",1,-1))

#tidyquant functions search for and key off a date or POSIXct column. 
#So, need to convert trial to that.
lr <- lr %>% mutate(trialDate = as_date(trial))
lr<- as_tibble(lr)
k <- 0:4
col_names <- paste0("lag_", k)

lags <- lr %>% group_by(subject) %>%
    tidyquant::tq_mutate(
        select     = orientCI,
        mutate_fun = lag.xts,
        k          = k,
        col_rename = col_names
    )
print(lags )

```

Calculate left bias for each trial
```{r}
lags<- lags %>% mutate(leftApproxCorr = abs(Left)<2, rightApproxCorr = abs(Right)<2 )
lags$leftApproxCorr <- as.numeric(lags$leftApproxCorr)
lags$rightApproxCorr <- as.numeric(lags$rightApproxCorr)

lags$leftBias <- lags$leftApproxCorr - lags$rightApproxCorr
```

Correlation of leftBias with orientation on this trial is pretty small (`r cor(lags$orientCI,lags$leftBias)` )

Calculate correlation with left performance instead of left bias
```{r}
cor(lags$orientCI, lags$leftApproxCorr)
```

Calculate autocorrelation

First, we need to correlate each of the lags to the “leftBias” column. This involves a few steps that can be strung together in a dplyr pipe (%>%):

The goal is to get orientCI and each lag side-by-side so we can do a correlation. To do this we use gather() to pivot each of the lagged columns into a “tidy” (long format) data frame, and we exclude “package”, “date”, and “count” columns from the pivot.

```{r}
# Calculate the autocorrelations and 95% cutoffs
lr_autocorr <- lags %>%
    gather(key = "lag", value = "lag_value", -c(subject, trialDate,leftBias, orientation,trial,Left,Right,leftApproxCorr,rightApproxCorr)) %>%
    mutate(lag = str_sub(lag, start = 5) %>% as.numeric) %>%
    group_by(subject, lag) %>%
    summarize(
        cor = cor(x = leftBias, y = lag_value, use = "pairwise.complete.obs"),
        cutoff_upper = 2/(n())^0.5,
        cutoff_lower = -2/(n())^0.5
        )
lr_autocorr
```


Now that we have the correlations calculated by package and lag number in a nice “tidy” format, we can visualize the autocorrelations

```{r , fig.height=5, fig.width=7}
library('papaja')
lr_autocorr %>%
    ggplot(aes(x = lag, y = cor, color = subject, group = subject)) +
    # Add horizontal line a y=0
    geom_hline(yintercept = 0) +
    # Plot autocorrelations
    geom_point(size = .5) +
      stat_summary(fun.y="mean",geom="line") + #,aes(group=orientation))
   # geom_segment(aes(xend = lag, yend = 0), size = 1) +
    # Plot summaries
stat_summary(fun.y="mean",geom="point",size=4, aes(group="abc"),color="black") + 
  stat_summary(fun.data = mean_cl_boot, geom="errorbar",  aes(group="abc"), width=0.4, color="black") +  
      stat_summary(fun.y="mean",geom="line",aes(group="abc"),color="black",size=1) + 
    # Add facets
    #facet_wrap(~ subject, ncol = 3) +
    # Aesthetics
    expand_limits(y = c(-.5, .5)) +
   theme(legend.position = "none") +
    #scale_color_tq() +
    xlab("Lag") + ylab('Correlation of leftBias with orientation')+
  ggtitle('Correlation of left bias with this trial orientation\n and previous trial orientations') + theme_apa()

ggsave(file.path("Results","Plots","autocorrE1.png"))
```

Use stat_summary to average across Ss in plot

```{r }
lr_autocorr %>% 
    ggplot(aes(x = lag, y = cor)) +
    # Add horizontal line a y=0
    geom_hline(yintercept = 0) +
    # Plot autocorrelations
stat_summary(fun.y="mean",geom="point",size=4) + 
  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4) +
    # Add facets
    #facet_wrap(~ subject, ncol = 3) +
    # Aesthetics
    expand_limits(y = c(-.1, .1)) +
  ggtitle('Autocorrelation') + 
    #scale_color_tq() +
    stat_summary(fun.y="mean",geom="line") + #,aes(group=orientation))
    theme_tq() +
    xlab("Lag") + ylab('Correlation of leftBias with orientation of previous trial')
    theme(
        legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)
    )
```

Collapse across subjects so can potentially report exact values of correlations.

```{r}
lr_autocorr %>% group_by(lag) %>% summarise(cor=mean(cor))
```

In case of nonlinearities, compare streaks of 3 in a row to everything else.

```{r}
lagss<- lags %>% mutate(streak5 = (lag_0==lag_1) & (lag_1==lag_2) & (lag_2==lag_3) & (lag_3==lag_4))
lagss<- lagss %>% mutate(streak4 = (lag_0==lag_1) & (lag_1==lag_2) & (lag_2==lag_3))
table(lagss$streak4)
table(lagss$streak5)
```

Plot left bias for streak versus not, grouped by orientation.

First collapse trials to calculate average within subject.
```{r, plotstreak}
lagavg <- lagss %>% filter(!is.na(streak4)) %>% 
                  group_by(subject,orientation,streak4) %>%
                summarise(leftBias = mean(leftBias))
```

```{r, plotstreak}
ggplot(lagss, aes(x=orientation,y=leftBias,color=streak4)) +
  stat_summary(fun.y="mean",geom="point") +
  theme_apa()
```

